### 2024-04-22 paper/paper.qmd
Update paper, llm usage, and references
Delete 05 replication script

---
title: "The Impact of the 2022-23 NBA Rule Change"
subtitle: "Exploring the Relationship Between the New Transitional Foul Rule, Player Shooting Performance, and MVP Points"
author: 
  - Amie Liu
thanks: "Code and data are available at: https://github.com/AmieLiu/the-NBA-Rule-Change.git."
date: today
date-format: long
abstract: "Rule changes in professional basketball play a critical role in shaping both strategies and player performances. Specifically, the introduction of the transitional foul rule in the 2022-23 NBA season, which changes the way fouls are called during transition plays, could significantly affect players' shooting efficiency and their prospects for MVP awards. This paper examines the implications of this rule change, comparing player statistics between the 2022-23 and 2021-22 seasons. We find significant correlations between shooting efficiency and MVP points, highlighting the rule's influence on player recognition. This paper offers insights into the interplay between rule modifications, player performance, and individual accolades, underscoring the broader significance of adapting to regulatory shifts in professional sports."
format: pdf
toc: true
number-sections: true
bibliography: references.bib
---

```{r}
#| include: false
#| warning: false
#| message: false

library(tidyverse)
library(knitr)
library(dplyr)
library(cowplot)
library(broom)

# Read data
analyzed_data2023 <- read_csv("../data/analysis_data/analysis_data2023.csv")
analyzed_data2022 <- read_csv("../data/analysis_data/analysis_data2022.csv")
```

# Introduction

In professional basketball, rule changes can reshape the game's landscape, altering strategies, player performances, and ultimately, the trajectory of teams' successes. The National Basketball Association (NBA), as the premier basketball league globally, often implements rule modifications to refine gameplay, increase scoring opportunities, and foster a more open and free-flowing style of basketball [@nba]. Under the new transitional foul rule in the 2022-23 NBA season, take fouls committed during transition scoring opportunities will result in one free throw and continued possession for the offended team [@rule], encouraging defenders to make legitimate plays.

This paper analyzes the consequences of the 2022-23 NBA rule change, particularly its impact on player shooting performance and the dynamics of the Most Valuable Player (MVP) competition, comparing statistics with the previous season, 2021-22. Our estimand is to explore the relationship between player shooting performance and MVP points they won in these NBA seasons. We hypothesize that players who excel in transition offense may experience an increase in shooting efficiency as they capitalize on more fast-break opportunities without facing as many transition-take fouls, potentially leading to greater recognition in MVP voting due to the positive correlation between shooting efficiency and MVP consideration.

A clear gap exists in understanding the nuanced impacts of the new transitional foul rule on player performance and MVP considerations, especially when compared to the previous season. This paper seeks to fill this gap by thoroughly analyzing the statistical data from both the 2022-23 and 2021-22 NBA seasons to highlight the changes, continuities, and emerging trends over these times. To address this gap, we used the R programming language [@citeR] to analyze shooting performance metrics, such as field goal percentage (FG%), three-point percentage (3P%), free throw percentage (FT%), as well as MVP voting points won from both seasons. Our findings reveal patterns in player shooting performance, MVP points accumulation, and the intertwined effects of the transitional foul rule… We discovered significant correlations between shooting efficiency and MVP points, with distinct variations observed between the two seasons.

Understanding the implications of the transitional foul rule in comparison to the previous season is crucial for stakeholders across the NBA landscape, from players and coaches to researchers and fans. By illuminating the intricate relationship between rule changes, player performance, and individual accolades, this paper contributes to a deeper understanding of the evolving dynamics within the league.

This paper is structured into three main sections: Data, Results, and Discussion. @sec-data Data explores sources and methodology. It includes various sources used for data collection, cleaning, and analyzing processes, drawing from the dataset. In @sec-results Results, we present our findings, focusing on data visualizations and analyses. @sec-discussion Discussion provides a comprehensive analysis of our research, highlights the implications for NBA basketball, the importance of limitations and considerations in our analysis, and identifies areas for future research. @sec-conclusion Conclusion summarizes our main findings and the broader significance of our research.

# Data {#sec-data}

The data for this paper are obtained from Basketball Reference [@basketball], which provides a wide range of basketball-related data that includes statistics, player and team information, game results, and more. The datasets used here are @basketball2023 and @basketball2022. These datasets include various statistics and information about NBA players who received votes for the MVP award, as well as their performance metrics for the 2022-23 and 2021-22 seasons. The data analysis involved using the R programming language [@citeR], along with the `tidyverse` [@citeTidyverse], `janitor` [@citeJanitor], `knitr` [@citeKnitr], `cowplot` [@cowplot], `broom` [@broom], and `rstanarm` [@rstanarm] packages for data collection, cleaning, analysis, and statistical modeling. Within the `tidyverse` package [@citeTidyverse], I used `readr` [@citeReadr], `dplyr` [@citeDplyr], `tidyr` [@citeTidyr], and `ggplot2` [@citeGgplot2] for reading CSV files, manipulating data, tidying data, and creating figures.

## NBA 2021-22 and 2022-23 Seasons Top 10 MVP Data

@tbl-table2023 presents the performance metrics of NBA players who received votes for the Most Valuable Player (MVP) award in the 2022-23 season. It includes a total of 10 rows and five variables: ranking of the top 10 players (Rank), the total number of MVP points won by each player (MVP Points Won), FG% (Field Goal Percentage), TP% (Three-Point Percentage), and FT% (Free Throw Percentage).

```{r}
#| label: tbl-table2023
#| tbl-cap: 2022-2023 season top 10 MVP table 
#| echo: false
#| warning: false

# Display the "analyzed_data2023" dataset
analyzed_data2023 |>
  kable(
    # Give each column a name
    col.names = c("Rank", "MVP Points Won", "Field Goal Percentage", "Three Points Percentage", "Free Throw Percentage")
  )
```

Field Goal Percentage indicates the efficiency of a player's shooting from the field, calculated as the ratio of successful field goals made to total field goal attempts. A higher field goal percentage suggests better shooting accuracy. Moreover, Three Points Percentage indicates the accuracy of a player's three-point shooting, measured as the percentage of successful three-point shots made out of total attempts. A higher three-point percentage reflects proficiency in long-range shooting. In addition, Free Throw Percentage represents the proficiency of a player in making free throws, calculated as the ratio of successful free throws made to total free throw attempts. A higher free throw percentage indicates better accuracy and consistency in free throw shooting.

@tbl-table2022 presents performance metrics of NBA players who received votes for the MVP award during the 2021-22 season. This dataset mirrors the structure of the previously mentioned dataset (see @tbl-table2023), providing insights into key performance indicators for players in the 2021-22 season.

```{r}
#| label: tbl-table2022
#| tbl-cap: 2021-2022 season top 10 MVP table
#| echo: false
#| warning: false

# Display the "analyzed_data2022" dataset
analyzed_data2022 |>
  kable(
    # Give each column a name
    col.names = c("Rank", "MVP Points Won", "Field Goal Percentage", "Three Points Percentage", "Free Throw Percentage")
  )
```

## Data Measurement

The measurement process quantifies real-world phenomena related to NBA player performance and translates them into dataset entries. We focused on performance metrics during the 2021-22 and 2022-23 seasons, including field goal percentage, three-point percentage, and free throw percentage. Data were collected from reputable sources like @basketball. These raw data were then processed and cleaned to ensure accuracy and consistency. The selected variables were defined and operationalized. For example, field goal percentage was calculated as the ratio of successful field goals made to attempts. Similarly, three-point and free throw percentages were computed using standardized formulas. These measurements became dataset entries, with rows representing players and columns representing performance metrics.

## Data Visualization
### The distribution of MVP points won by top 10 players in the 2021-22 and 2022-23 seasons

@fig-points illustrates the MVP points won by the top 10 NBA players in each of the 2022-23 and 2021-22 seasons. It clearly shows that the highest number of MVP points, which totaled 915, was given to the player who ranked 1st in the 2022-23 season. In contrast, the minimum point (1 point) was given to the player who ranked 10th in the 2021-22 season.

When examining the MVP points across the corresponding ranks in both seasons, players in seven out of the ten ranks during the 2022-23 season received more points than their counterparts in the same ranks during the 2021-22 season. The top ranks in the 2022-23 season (Rank 1, 3, 4) generally received more points than in the previous season. This might indicate a shift in voting behavior or a stronger consensus among voters regarding the leading candidates. Players at lower ranks (6-10) showed a relatively small variation in points, suggesting that while the top contenders might shift significantly in performance or perception, those at lower ranks maintain a more consistent level of recognition.

```{r}
#| label: fig-points
#| fig-cap: Number of MVP points won by each player from the top 10 rankings in the 2021-22 and 2022-23 seasons
#| fig.width: 8 
#| echo: false
#| warning: false
#| message: false

# Create a bar chart to visualize the distribution of MVP points won by rank based on the analyzed_data2023 and analyzed_data2022 datasets

# Add a "Season" column to the 2022-23 dataset
analyzed_data2023 <- analyzed_data2023 |>
  mutate(season = "2022-23")

# Add a "Season" column to the 2021-22 dataset
analyzed_data2022 <- analyzed_data2022 |>
  mutate(season = "2021-22")

# Combine the 2022-23 and 2021-22 datasets
combined_data <- bind_rows(analyzed_data2023, analyzed_data2022)

# Create the bar plot with rank on the x-axis and season as fill color
combined_data |>
  ggplot(aes(x = factor(rank), fill = factor(season))) +
  # Add a grouped bar layer to create the grouped bar chart
  geom_bar(position = "dodge", aes(y = points_won), stat = "identity") +
  # Add text labels for the number of points won on top of each bar
  geom_text(aes(label = points_won, y = points_won), vjust = -0.5, size = 3, position = position_dodge(width = 0.9)) +
  # Scale the x-axis labels to display "Rank"
  scale_x_discrete("Rank") +
  # Scale the fill color to display "season" and specify the break points: 2021-22 Season", "2022-23 Season"
  scale_fill_discrete(name = "Season",
                      breaks = c("2021-22", "2022-23"),
                      labels = c("2021-22 Season", "2022-23 Season")) +
  # Apply a minimal theme to the chart
  theme_minimal() +
  # Label y-axis as "Number of MVP Points Won"
  labs(y = "Number of MVP Points Won")
```

### Different shooting performances by top 10 players in the 2021-22 and 2022-23 seasons

@fig-percentage provides a comparison of the shooting performances of the top 10 NBA players across 2022-23 and 2021-22 seasons, including Field Goal, Three-Point, and Free Throw percentages. Each metric is depicted with trend lines to illustrate the general trends across player ranks from 1 to 10.

The highest Field Goal Percentage observed was 0.632 in the 2022-23 season. In contrast, the lowest percentage was recorded at 0.437 in the 2021-22 season. The trend lines show that the 2022-23 season has higher percentages across the ranks compared to the previous season, despite a slight decline in percentage from the highest to the lowest ranked players in both seasons. This means there is an enhancement in field goal accuracy from the 2021-22 season to the 2022-23 season.

For the Three-Point Percentage, the highest was recorded at 0.427 in the 2022-23 season, while the lowest was also observed in the same season at 0.275. Interestingly, the trend line for 2022-23 shows a slight increase as player rank increases, unlike in 2021-22, where the trend line remains relatively flat across ranks. There is a marginal increase in three-point shooting percentages in the 2022-23 season

The scatter plot for Free Throw Percentage shows the highest performance at 0.923 in the 2021-22 season and the lowest at 0.645 in the 2022-23 season. Both seasons exhibit a slight upward trend in free throw efficiency from rank 1 to 10. The free throw percentage remains nearly constant in both seasons, with a minimal increase in the 2021-22 season.

```{r}
#| echo: false
#| warning: false
#| message: false

# Create the scatter plot with trend line or smoothed line for "field_goal_percentage"
scatterplot1 <- 
  ggplot(combined_data, aes(x = rank, y = field_goal_percentage, color = season)) +
  geom_point() +  # Add points for each data point
  geom_smooth(method = "lm", se = FALSE) +  # Add a trend line (linear regression)
  # Scale the x-axis labels to display "Rank"
  scale_x_continuous(breaks = 1:10, labels = 1:10) +
  # Scale the y-axis labels to display "Field Goal Percentage"
  scale_y_continuous("Field Goal Percentage") +
  # Change legend title, position, and x-axis
  labs(title = "Field Goal Percentage by Rank", color = "Season", x = "Rank") +
  theme_minimal()
```

```{r}
#| echo: false
#| warning: false
#| message: false

# Create the scatter plot with trend line or smoothed line for "three_points_percentage"
scatterplot2 <- 
  ggplot(combined_data, aes(x = rank, y = three_points_percentage, color = season)) +
  geom_point() +  # Add points for each data point
  geom_smooth(method = "lm", se = FALSE) +  # Add a trend line (linear regression)
  # Scale the x-axis labels to display "Rank"
  scale_x_continuous(breaks = 1:10, labels = 1:10) +
  # Scale the y-axis labels to display "Three Points Percentage"
  scale_y_continuous("Three Points Percentage") +
  # Change legend title,  position, and x-axis
  labs(title = "Three Points Percentage by Rank", color = "Season", x = "Rank") +
  theme_minimal()
```

```{r}
#| label: fig-percentage
#| fig-cap: "Different shooting percentage by each player from the top 10 rankings in the 2021-22 and 2022-23 seasons"
#| fig.height: 10
#| fig.width: 8 
#| echo: false
#| warning: false
#| message: false

# Create the scatter plot with trend line or smoothed line for "free_throw_percentage"
scatterplot3 <-
  ggplot(combined_data, aes(x = rank, y = free_throw_percentage, color = season)) +
  geom_point() +  # Add points for each data point
  geom_smooth(method = "lm", se = FALSE) +  # Add a trend line (linear regression)
  # Scale the x-axis labels to display "Rank"
  scale_x_continuous(breaks = 1:10, labels = 1:10) +
  # Scale the y-axis labels to display "Free Throw Percentage"
  scale_y_continuous("Free Throw Percentage") +
  # Change legend title, position, and x-axis
  labs(title = "Free Throw Percentage by Rank", color = "Season", x = "Rank") +
  theme_minimal()

# Create the combined plot
combined_plot <- plot_grid(scatterplot1, scatterplot2, scatterplot3, ncol = 1)

combined_plot
```

# Model {#sec-model}

To explore the impact of the 2022-23 NBA transitional foul rule on player shooting performance and its subsequent effect on MVP voting, we constructed a linear regression model. The objective is to quantify the relationship between shooting performance metrics and MVP voting outcomes.

## Model set-up

The dependent variable is MVP Voting Points and the independent variables are Field Goal Percentage (FG%), Three-Point Field Goal Percentage (3P%), and Free Throw Percentage (FT%). Each independent variable was chosen based on its potential influence on a player's performance under the conditions created by the new foul rule. We run the model in R [@citeR] utilizing the `rstanarm` package [@rstanarm]. Our model is constructed as follows:

$$
Y = \beta_0 + \beta_1X_1 + \beta_2X_2 + \beta_3X_3 + \epsilon
$$

where:

- $Y$ is the MVP Voting Points. It is the dependent variable that represents the points a player receives in the MVP voting process.
- $\beta_0$ is the intercept of the regression model that represents the expected value of Total MVP Voting Points when all predictor variables (FG%, 3P%, FT%) are at zero.
- $\beta_1$ is the coefficient for FG%. It indicates how changes in a player’s overall shooting efficiency from the field affect their MVP voting points.
- $\beta_2$ is the the coefficient for 3P%. This reflects the impact of three-point shooting efficiency on MVP voting.
- $\beta_3$ is the coefficient for FT%. This measures the influence of free throw shooting efficiency on MVP voting points.
- $X_1$ measures the FG% made out of attempts.
- $X_2$ measures the 3P% made out of attempts.
- $X_3$ measures the FT% made out of attempts.
- $\epsilon$ is the error term, representing the residual element not explained by the model. This captures the variability in MVP voting points not explained by shooting percentages alone.

### Model justification

The selection of linear regression for our analysis suggests linear relationships between the shooting metrics and MVP voting points. We hypothesized that improvements in these shooting metrics, reflecting enhanced performance in transition plays free from take fouls, would correlate positively with higher MVP voting scores. The use of FG%, 3P%, and FT% as predictors is justified by their direct impact on game performance, which is critical in MVP considerations. Moreover, these metrics are commonly used in basketball analytics to assess player efficiency and effectiveness, making them relevant for evaluating the consequences of rule changes.

Statistical significance of the model coefficients will be tested using t-tests, where a p-value less than 0.05 will indicate significant influence of the predictor on MVP voting points. The overall fit of the model will be assessed by the R-squared value, which explains the proportion of variance in MVP points accounted for by the predictors.

# Results {#sec-results}

Our results are summarized in @tbl-coefficients and @tbl-additional (see @fig-modelresults)

```{r}
#| echo: false
#| eval: true
#| warning: false
#| message: false

model2023 <-
  readRDS(file = "../models/model2023.rds")
```

```{r}
#| label: tbl-coefficients
#| tbl-cap: "Summary of regression model coefficients"
#| echo: false
#| eval: true
#| warning: false
#| message: false

# Generate a tidy summary of the model
model_summary <- tidy(model2023) |>
  rename(
    Term = term,
    Estimate = estimate,
    `Std. Error` = std.error,
    Statistic = statistic,
    `P-value` = p.value,
  ) |>
  # Rename the "Term" values
  mutate(
    Term =
      case_match(
        Term,
        "(Intercept)" ~ "(Intercept)",
        "field_goal_percentage" ~ "Field Goal Percentage",
        "three_points_percentage" ~ "Three Points Percentage",
        "free_throw_percentage" ~ "Free Throw Percentage"
        )
    )

# Use knitr's kable function to format tables
kable(model_summary)
```

```{r}
#| label: tbl-additional
#| tbl-cap: "Summary of additional regression model metrics"
#| echo: false
#| eval: true
#| warning: false
#| message: false

# Include additional model metrics
additional_metrics <- data.frame(
  Metric = c("Residual Standard Error", "Multiple R-squared", "Adjusted R-squared", "F-statistic", "p-value"),
  Value = c("314.3", "0.4446", "0.1669", "1.601", "0.2852")
)

kable(additional_metrics)
```

```{r}
#| label: fig-modelresults
#| fig-cap: "Observed vs. predicted values of MVP points won"
#| fig.width: 8 
#| echo: false
#| warning: false
#| message: false

# Diagnostic plots
#par(mfrow = c(2, 2))  # Arrange plots in a 2x2 grid
#plot(model2023)  # Generate plots

# Create a data frame of observed vs. predicted values
predicted_values <- data.frame(Observed = analyzed_data2023$points_won, 
                               Predicted = predict(model2023))

# Plot observed vs. predicted values
observed_predicted_values <-
  ggplot(predicted_values, aes(x = Observed, y = Predicted)) +
  geom_point() +
  geom_abline(color = "red", linetype = "dashed") +  # Line x=y for reference
  labs(x = "Observed Points",
       y = "Predicted Points") +
  theme_minimal()

observed_predicted_values
```

# Discussion {#sec-discussion}

## First discussion point {#sec-first-point}

If my paper were 10 pages, then should be be at least 2.5 pages. The discussion is a chance to show off what you know and what you learnt from all this.

## Second discussion point

## Third discussion point

## Weaknesses and next steps

Weaknesses and next steps should also be included.

\newpage

# Conclusion {#sec-conclusion}

\newpage

\appendix
# Appendix {-}

# Additional data details

# Model details {#sec-model-details}

## Posterior predictive check

In @fig-ppcheckandposteriorvsprior-1 we implement a posterior predictive check. This shows...

In @fig-ppcheckandposteriorvsprior-2 we compare the posterior with the prior. This shows...

\newpage

# References